Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-02-09 Manuscript/Tmatrix.tex
data availibility statement
significant modifications in the appendix

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%
The $GW$ approximation \cite{Hedin_1965} of many-body perturbation theory \cite{Martin_2016} is becoming a method of choice to target charged excitations (\ie, ionization potentials and electron affinities) in molecular systems. \cite{Aryasetiawan_1998,Onida_2002,Reining_2017,Golze_2019,Bruneval_2021} 
These so-called quasiparticle energies can be experimentally measured from direct and inverse photoemission spectroscopies.
From a more theoretical point of view, $GW$ corresponds to an elegant resummation of all direct ring diagrams from the particle-hole (ph) channel which is particularly justified in the high-density or weakly-correlated regime. \cite{Gell-Mann_1957,Nozieres_1958}
Within the $GW$ approximation, the self-energy --- one of the key quantities of Hedin's equations \cite{Hedin_1965} --- reads
\begin{equation}
	\Sigma^{GW}(1,2) = i G(1,2) W(1,2)
\end{equation}
where $G$ is the one-body Green's function, $W$ is the dynamically-screened Coulomb potential, and, \eg, $1 \equiv (\sigma_1,\br_1,t_1)$ is a composite coordinate gathering spin, space, and time variables.

Alternatives to $GW$ do exist. For example, the $T$-matrix (or Bethe-Goldstone) approximation, first introduced in nuclear physics, \cite{Bethe_1957,Baym_1961,Baym_1962,Danielewicz_1984a,Danielewicz_1984b} then in condensed matter physics, \cite{Liebsch_1981,Bickers_1989,Bickers_1991,Katsnelson_1999,Katsnelson_2002,Zhukov_2005,vonFriesen_2010,Romaniello_2012,Gukelberger_2015} and more recently in quantum chemistry, \cite{Zhang_2017,Li_2021b} sums to infinity the ladder diagrams from the particle-particle (pp) channel and is justified in the low-density or strongly-correlated regime. \cite{Danielewicz_1984a,Danielewicz_1984b,Liebsch_1981,Shepherd_2014}
While the two-point screened interaction $W$ is the cornerstone of $GW$, the $T$-matrix approximation relies on a more complex (four-point) effective interaction --- the so-called $T$ matrix --- yielding the following self-energy: 
\begin{equation}
	\Sigma^{GT}(1,2) = i \int G(4,3) T(1,3;2,4) d3 d4
\end{equation}
The natural idea of combining the ph and pp channels is also possible and has been explored, for example, in the Hubbard dimer within many-body perturbation theory \cite{Romaniello_2012} and the uniform electron gas \cite{Loos_2016} within coupled-cluster theory. \cite{Shepherd_2014}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Neutral excitations}
\label{sec:neutral}
%%%%%%%%%%%%%%%%%%%%%%
Like the one-body Green's function is the pillar of the $GW$ and $GT$ approximations, the two-body Green's function $G_2$ is the key quantity of the Bethe-Salpeter equation (BSE) formalism of many-body perturbation theory \cite{Salpeter_1951,Strinati_1988,Blase_2018,Blase_2020} via its link with the two-body correlation function $L$ which satisfies the self-consistent BSE 
\begin{multline}
	 iL(1,2;1',2') = L_0(1,2;1',2') 
	 \\
	 + \int L_0(1,4;1',3) \Xi(3,5;4,6) L(6,2;5,2') d3d4d5d6
\end{multline}
where
\begin{subequations}
\begin{align}
	 iL_0(1,4;1',3) & = G(1,3) G(4,1')
	 \\
	 iL(1,2;1',2') & = -G_2(1,2;1',2') + G(1,1')G(2,2')
\end{align}
\end{subequations}
and 
\begin{equation}
\label{eq:Xi}
	\Xi(3,5;4,6) = i \fdv{\Sigma(3,4)}{G(6,5)}
\end{equation}
is the so-called BSE kernel that takes into account the self-consistent variation of $\Sigma$ with respect to the variation of $G$.
By taking into account the interaction of the excited electron and its hole left behind (the infamous excitonic effect), the BSE is able to faithfully model (neutral) optical excitations as measured by absorption spectroscopy. 
The moderate cost of the BSE [which scales as $\order*{K^4}$ in its standard implementation] and its all-round accuracy are the main reasons behind its growing popularity in the molecular electronic structure community. \cite{Rohlfing_1999a,Horst_1999,Puschnig_2002,Tiago_2003,Boulanger_2014,Jacquemin_2015a,Bruneval_2015,Jacquemin_2015b,Hirose_2015,Jacquemin_2017a,Jacquemin_2017b,Rangel_2017,Krause_2017,Gui_2018,Blase_2018,Liu_2020,Blase_2020,Holzer_2018a,Holzer_2018b,Loos_2020e,Loos_2021}

@@

To recast the BSE with a dynamical $T$-matrix kernel as an eigenvalue equation, we follow Ref.~\onlinecite{Strinati_1980} and start from the equation for the BSE amplitude 
\begin{equation}
\label{eq:Strinati_1}
	X_S(1,1^\prime) = \int d3d4d5d6L_0(1,4,1^\prime,3)\Xi(3,5,4,6) X_S(6,5)
\end{equation}
with 
\begin{align}
\label{eq:T-matrix_kernel}
	L_0(1,4,1^\prime,3) & =-iG(1,3)G(4,1^\prime)
	&
	\Xi(3,5,4,6) & =  i\frac{\delta \Sigma(3,4)}{\delta G(6,5)}\approx -T(3,5,4,6)
\end{align}
where we consider the so-called $T$-matrix self-energy \cite{Martin_2016,Romaniello_2012} $\Sigma(3,4)=iG(7,8)T(3,8,4,7)$ with
\begin{equation}
\label{eq:T-matrix}
	T(3,8,4,7) =
	- v_c(3,8)\delta(3,4)\delta(7,8)
	+ v_c(3,8)\delta(4,8)\delta(3,7)
	+ iv_c(3,8)G(3,1')G(8,2')T(1',2',4,7)
\end{equation}  
and where we neglect the functional derivative $\delta T/\delta G$ in the kernel $\Xi$.
Making the time dependence of Eq.~\eqref{eq:Strinati_1} explicit, one gets 
\begin{multline}
\label{eqn:DYN_T}
	X_S(x_1,x_{1^\prime},\tau_{11'})e^{-i\omega_S(t_1+t_{1^\prime})/2}
	=-i\int d3d4d5d6 G(x_1,x_3;\tau_{13})G(x_4,x_{1^\prime};\tau_{11^\prime})
	\\
	\times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{multline}
where $\tau^+_{ij}=t^+_i-t_j$ with $t^+_i=t_i+\eta$ ($\eta \to 0^+$) and $T(3,5,4,6)=- \delta(\tau^+_{35})\delta(\tau^+_{64})\bar{T}(x_3,x_5,x_4,x_6;\tau_{34})$ with 
\begin{multline}
\label{eq:T_bar}
	\bar{T}(x_3,x_5,x_4,x_6;\tau_{34}) 
	= v_c(x_3,x_5)\delta(\tau_{34})\delta(x_3,x_4)\delta(x_6,x_5)-v_c(x_3,x_5)\delta(\tau_{34})	\delta(x_3,x_6)\delta(x_4,x_5)
	\\
	+ iv_c(x_3,x_5)G(x_3,x_7;\tau_{38})G(x_5,x_8;\tau^+_{38})\bar{T}(x_7,x_8,x_4,x_6;\tau_{84})
\end{multline}
Using the Fourier transform $G(\tau)=\int \frac{d\omega}{2\pi}G(\omega)e^{-i\omega\tau}$, changing variable from $t_3$ to $\tau_{34}$ and taking the limit $t_{1^\prime}=t_1^+$ we arrive at
\begin{multline}
	X_S(x_1,x_{1^\prime},0^-)
	= -i\int dx_3 dx_4 dx_5 dx_6\int d\tau_{34} \int \frac{d\nu'}{2\pi}G(x_1,x_3;\nu'+\omega_S)G(x_4,x_{1^\prime};\nu')e^{i\nu'\tau_{34}}
	\\
	 \times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}.
\end{multline}
Using the Lehman representation of the one-body Green's function in the quasiparticle approximation, i.e.
$$G(x_1,x_2;\omega)=\sum_{p}\frac{\psi_p(x_1)\psi_p^*(x_2)}{\omega-\epsilon_p+ i\eta \text{sgn}(\epsilon_p-\mu)},$$ where $\psi_p$ and $\epsilon_p$ are single-particle wavefunctions and energies, respectively, and $\mu$ the chemical potential, and multiply left- and right-hand sides by $(\epsilon_a-\epsilon_i-\omega_S)\int dx_1dx^{\prime}_1 \psi^*_a(x_1)\psi_i(x_1^\prime)$.  
Using the fact that
\begin{equation}
	\Theta(\pm\tau)e^{-i\alpha \tau}=\mp\frac{1}{2\pi i}\lim_{\eta\to 0^+}\int d\omega \frac{1}{\omega-\alpha\pm i\eta}e^{-i\omega\tau}
\end{equation}
we then arrive at
\begin{multline}
\label{eq:General_DYN}
	(\epsilon_a-\epsilon_i-\omega_S)\int dx_1dx^{\prime}_1 \psi^*_a(x_1)\psi_i(x_1^\prime)  X_S(x_1,x_{1^\prime},0^-) =
	\\
	-\int dx_3 dx_4 dx_5 dx_6\int d\tau_{34}\psi_a^*(x_3)\psi_i(x_4)
	\qty[\Theta(\tau_{34})e^{i\epsilon_{i}\tau_{34}}+\Theta(-\tau_{34})e^{i(\epsilon_{a}-\omega_S)\tau_{34}} ]
	\\
	\times \bar{T}(x_3,x_5,x_4,x_6;\tau_{34})X_S(x_6,x_5;-\tau_{34})e^{i\omega_S\tau_{34}/2}
\end{multline}

We now consider the resonant case $\omega_m>0$, for which 
\begin{equation}
\label{Eqn:chi_tau}
	X_S(x_1,x_{1^\prime},\tau_1) 
	=-e^{i\omega_S\abs{\tau_1}/2} \sum_{pq}\psi_p(x_1)\psi^*_{q}(x_{1^\prime}) \mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
  	\qty[\Theta(\tau_1)e^{-i\epsilon_p\tau_1}+\Theta(-\tau_1)e^{-i\epsilon_{q}\tau_1} ]
\end{equation}
  where $\hat{c}^\dagger_{q}$ and $\hat{c}_p$ are creation and annihilation operators, respectively, and $\ket{N}$ and $\ket{N,S}$ are the ground state and the $S$-th excited state, respectively, of the $N$-electron system.
After some algebraic steps we arrive at 
\begin{multline}
	-(\epsilon_a-\epsilon_i-\omega_S)\mel{N}{\hat{c}^\dagger_{i}\hat{c}_a}{N,S}
	\\
    = -\sum_{pq}\mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
    \qty{ -\frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\ \bar{T}_{iq,pa}(\omega) e^{-i\omega\eta}\left[\frac{1}{\omega_S-\omega+\epsilon_{q}+\epsilon_i+i\eta}+\frac{1}{\omega_S+\omega-\epsilon_p-\epsilon_a+i\eta}\right] }
\end{multline}
where we define 
\begin{equation}
	\bar{T}_{iq,pa}(\tau_{34})=\int dx_3x_4x_5x_6\int \psi_a^*(x_3)\psi_i(x_4)\bar{T}(x_3,x_5,x_4,x_6;\tau_{34}) \psi_p(x_6)\psi^*_{q}(x_5)
\end{equation}
Using the definition $A_S^{(qp)}=\langle N|\hat{c}^\dagger_{q}\hat{c}_p|N,S\rangle$, we arrive at

\begin{equation}
	\qty(\epsilon_a-\epsilon_i-\omega_S)A_S^{(ia)}+\sum_{pq}A_S^{(q p)}\widetilde{T}_{iq,pa}(\omega_S)=0,
\end{equation}
where we defined
\begin{equation}
	\widetilde{T}_{iq,pa}(\omega_S)
	= \qty{ \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\bar{T}_{iq,pa}(\omega)e^{-i\omega\eta} \left[\frac{1}{\omega_S-\omega+\epsilon_{q}+\epsilon_i+i\eta}+\frac{1}{\omega_S+\omega-\epsilon_p-\epsilon_a+i\eta}\right]  }
 \label{Eqn:H2preso_eigen_2_T}
\end{equation}

This is an eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as,
\begin{equation}
\label{eq:H2preso_eigen}
	H^{2p,reso}_{ia,qp}(\omega_S)A_S^{(qp)}=\omega_S A^{(ia)}_S,
\end{equation}
with
\begin{equation}
	H^{2p,reso}_{ia,qp}(\omega_S)=(\epsilon_a-\epsilon_i)\delta_{iq}\delta_{ap} + \mel{ip}{}{aq}+\widetilde{T}_{iq,pa}(\omega_S).
\end{equation}
Here $H^{2p,reso}_{ia,jb}(\omega_S)=A_{ia,jb}(\omega_S)$ and $H^{2p,reso}_{ia,bj}(\omega_S)=B_{ia,jb}(\omega_S)$.
If we drop the correlation part of $\bar{T}$ [the first two terms on the right-hand side of Eq.~\eqref{eq:T_bar}] we obtain the TDHF equations. To calculate the correlation contribution we use \alert{Eq.~(9)} in Eq.~\eqref{Eqn:H2preso_eigen_2_T}, and after integration over the frequency we arrive at \alert{Eq.~(21).}

%%%%%%%%%%%%%%%%%%%%%%%%
\twocolumngrid
\bibliography{Tmatrix}
%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\DeclareSIUnit\angstrom{\text{\AA}}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%
The $GW$ approximation \cite{Hedin_1965} of many-body perturbation theory \cite{Martin_2016} is becoming a method of choice to target charged excitations (\ie, ionization potentials and electron affinities) in molecular systems. \cite{Aryasetiawan_1998,Onida_2002,Reining_2017,Golze_2019,Bruneval_2021} 
These so-called quasiparticle energies can be experimentally measured from direct and inverse photoemission spectroscopies.
From a more theoretical point of view, $GW$ corresponds to an elegant resummation of all direct ring diagrams from the particle-hole (ph) channel which is particularly justified in the high-density or weakly-correlated regime. \cite{Gell-Mann_1957,Nozieres_1958}
Within the $GW$ approximation, the self-energy --- one of the key quantities of Hedin's equations \cite{Hedin_1965} --- reads
\begin{equation}
\label{eq:SigGW}
	\Sigma^{GW}(1,2) = i G(1,2) W(1,2)
\end{equation}
where $G$ is the one-body Green's function, $W$ is the dynamically-screened Coulomb potential, and, \eg, $1 \equiv (\sigma_1,\br_1,t_1)$ is a composite coordinate gathering spin, space, and time variables.

Alternatives to $GW$ do exist. For example, the $T$-matrix (or Bethe-Goldstone) approximation, first introduced in nuclear physics, \cite{Bethe_1957,Baym_1961,Baym_1962,Danielewicz_1984a,Danielewicz_1984b} then in condensed matter physics, \cite{Liebsch_1981,Bickers_1989,Bickers_1991,Katsnelson_1999,Katsnelson_2002,Zhukov_2005,vonFriesen_2010,Romaniello_2012,Gukelberger_2015} and more recently in quantum chemistry, \cite{Zhang_2017,Li_2021b} sums to infinity the ladder diagrams from the particle-particle (pp) channel and is justified in the low-density or strongly-correlated regime. \cite{Danielewicz_1984a,Danielewicz_1984b,Liebsch_1981,Shepherd_2014}
While the two-point screened interaction $W$ is the cornerstone of $GW$, the $T$-matrix approximation relies on a more complex (four-point) effective interaction --- the so-called $T$ matrix --- yielding the following self-energy: 
\begin{equation}
\label{eq:SigGT}
	\Sigma^{GT}(1,2) = i \int G(4,3) T(1,3;2,4) d3 d4
\end{equation}
The natural idea of combining the ph and pp channels is also possible and has been explored, for example, in the Hubbard dimer within many-body perturbation theory \cite{Romaniello_2012} and the uniform electron gas \cite{Loos_2016} within coupled-cluster theory. \cite{Shepherd_2014}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Neutral excitations}
\label{sec:neutral}
%%%%%%%%%%%%%%%%%%%%%%
Like the one-body Green's function is the pillar of the $GW$ and $GT$ approximations, the two-body Green's function $G_2$ is the key quantity of the Bethe-Salpeter equation (BSE) formalism of many-body perturbation theory \cite{Salpeter_1951,Strinati_1988,Blase_2018,Blase_2020} via its link with the two-body correlation function $L$ which satisfies the self-consistent BSE 
\begin{multline}
	 iL(1,2;1',2') = L_0(1,2;1',2') 
	 \\
	 + \int L_0(1,4;1',3) \Xi(3,5;4,6) L(6,2;5,2') d3d4d5d6
\end{multline}
where
\begin{subequations}
\begin{align}
\label{eq:L0}
	 iL_0(1,4;1',3) & = G(1,3) G(4,1')
	 \\
\label{eq:L}
	 iL(1,2;1',2') & = -G_2(1,2;1',2') + G(1,1')G(2,2')
\end{align}
\end{subequations}
and 
\begin{equation}
\label{eq:Xi}
	\Xi(3,5;4,6) = i \fdv{\Sigma(3,4)}{G(6,5)}
\end{equation}
is the so-called BSE kernel that takes into account the self-consistent variation of $\Sigma$ with respect to the variation of $G$.
By taking into account the interaction of the excited electron and its hole left behind (the infamous excitonic effect), the BSE is able to faithfully model (neutral) optical excitations as measured by absorption spectroscopy. 
The moderate cost of the BSE [which scales as $\order*{K^4}$ in its standard implementation] and its all-round accuracy are the main reasons behind its growing popularity in the molecular electronic structure community. \cite{Rohlfing_1999a,Horst_1999,Puschnig_2002,Tiago_2003,Boulanger_2014,Jacquemin_2015a,Bruneval_2015,Jacquemin_2015b,Hirose_2015,Jacquemin_2017a,Jacquemin_2017b,Rangel_2017,Krause_2017,Gui_2018,Blase_2018,Liu_2020,Blase_2020,Holzer_2018a,Holzer_2018b,Loos_2020e,Loos_2021}

@@

To recast the BSE with a dynamical $T$-matrix kernel as an eigenvalue equation, we follow Ref.~\onlinecite{Strinati_1980} (see also Ref.~\onlinecite{Loos_2020e}) and start from the equation for the BSE amplitude 
\begin{equation}
\label{eq:Strinati_1}
	\chi_S(1,1') = \int d3d4d5d6L_0(1,4,1',3)\Xi(3,5,4,6) \chi_S(6,5)
\end{equation}
where $L_0$ is given by Eq.~\eqref{eq:L0} and the $T$-matrix kernel is
\begin{align}
\label{eq:T-matrix_kernel}
	\Xi(3,5,4,6) & =  i\frac{\delta \Sigma(3,4)}{\delta G(6,5)}\approx -T(3,5,4,6)
\end{align}
The so-called $T$-matrix self-energy $\Sigma$ is given by Eq.~\eqref{eq:SigGT} with \cite{Martin_2016,Romaniello_2012}
\begin{equation}
\label{eq:T-matrix}
	T(3,8,4,7) =
	- v_c(3,8)\delta(3,4)\delta(7,8)
	+ v_c(3,8)\delta(4,8)\delta(3,7)
	+ i \int d1'd2' v_c(3,8)G(3,1')G(8,2')T(1',2',4,7)
\end{equation}  
where \titou{$v_c$ is the bare Coulomb operator} and we neglect the functional derivative $\delta T/\delta G$ in the kernel $\Xi$.
Making the time dependence of Eq.~\eqref{eq:Strinati_1} explicit, one gets 
\begin{multline}
\label{eqn:DYN_T}
	\chi_S(\bx_1,\bx_{1'},\tau_{11'})e^{-i\omega_S(t_1+t_{1'})/2}
	=-i\int d3d4d5d6 G(\bx_1,\bx_3;\tau_{13})G(\bx_4,\bx_{1'};\tau_{11'})
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{multline}
[where \titou{$\tau_{ij}=t_i-t_j$} and $\tau^+_{ij}=t^+_i-t_j$ with $t^+_i=t_i+\eta$ ($\eta \to 0^+$)] with $T(3,5,4,6)=- \delta(\tau^+_{35})\delta(\tau^+_{64})\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})$ and
\begin{multline}
\label{eq:T_bar}
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34}) 
	= v_c(\bx_3,\bx_5)\delta(\tau_{34})\delta(\bx_3,\bx_4)\delta(\bx_6,\bx_5)-v_c(\bx_3,\bx_5)\delta(\tau_{34})	\delta(\bx_3,\bx_6)\delta(\bx_4,\bx_5)
	\\
	+ iv_c(\bx_3,x_5)G(\bx_3,\bx_7;\tau_{38})G(\bx_5,\bx_8;\tau^+_{38})\Bar{T}(\bx_7,\bx_8,\bx_4,\bx_6;\tau_{84})
\end{multline}
Using the Fourier transform $G(\tau)=\int \frac{d\omega}{2\pi}G(\omega)e^{-i\omega\tau}$, changing variable from $t_3$ to $\tau_{34}$ and taking the limit $t_{1'}=t_1^+$, we have
\begin{multline}
	\chi_S(\bx_1,\bx_{1'},0^-)
	= -i\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34} \int \frac{d\omega'}{2\pi}G(\bx_1,\bx_3;\omega'+\omega_S)G(\bx_4,\bx_{1'};\omega')e^{i\omega'\tau_{34}}
	\\
	 \times \Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{multline}
Using the Lehman representation of the one-body Green's function in the quasiparticle approximation given by Eq.~\eqref{eq:G},
%$$G(x_1,x_2;\omega)=\sum_{p}\frac{\psi_p(x_1)\psi_p^*(x_2)}{\omega-\epsilon_p+ i\eta \text{sgn}(\epsilon_p-\mu)},$$ where $\psi_p$ and $\epsilon_p$ are single-particle wavefunctions and energies, respectively, and $\mu$ the chemical potential, 
and multiply the left- and right-hand sides by $(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1 d\bx^{\prime}_1 \SO{a}(\bx_1)\SO{i}(\bx_1')$.  
we then arrive at
\begin{multline}
\label{eq:General_DYN}
	\qty(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1d\bx^{\prime}_1 \psi^*_a(\bx_1)\psi_i(\bx_1')  \chi_S(\bx_1,\bx_{1'},0^-) =
	\\
	-\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34}\psi_a(\bx_3)\psi_i(\bx_4)
	\qty[\Theta(\tau_{34})e^{i\epsilon_{i}\tau_{34}}+\Theta(-\tau_{34})e^{i(\epsilon_{a}-\omega_S)\tau_{34}} ]
	\\
	\times \Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{i\omega_S\tau_{34}/2}
\end{multline}
(where $\Theta$ is the Heaviside step function) using the fact that
\begin{equation}
	\Theta(\pm\tau)e^{-i\alpha \tau}=\mp\frac{1}{2\pi i}\lim_{\eta\to 0^+}\int d\omega \frac{1}{\omega-\alpha\pm i\eta}e^{-i\omega\tau}
\end{equation}

We now consider the resonant case $\omega_S>0$, for which 
\begin{equation}
\label{Eqn:chi_tau}
	\chi_S(\bx_1,\bx_{1'},\tau_1) 
	=-e^{i\omega_S\abs{\tau_1}/2} \sum_{pq}\SO{p}(\bx_1)\SO{q}(\bx_{1'}) \mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
  	\qty[\Theta(\tau_1)e^{-i\e{p}{}\tau_1}+\Theta(-\tau_1)e^{-i\e{q}{}\tau_1} ]
\end{equation}
  where $\hat{c}^\dagger_{q}$ and $\hat{c}_p$ are creation and annihilation operators, respectively, and $\ket{N}$ and $\ket{N,S}$ are the ground state and the $S$-th excited state, respectively, of the $N$-electron system.
After some algebraic steps, one gets
\begin{multline}
	-\qty(\e{a}{}-\e{i}{}-\omega_S)\mel{N}{\hat{c}^\dagger_{i}\hat{c}_a}{N,S}
	\\
    = -\sum_{pq}\mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}
    \qty{ -\frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\ \Bar{T}_{iq,pa}(\omega) e^{-i\omega\eta} \qty[\frac{1}{\omega_S-\omega+\e{q}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{p}{}-\e{a}{}+i\eta} ] }
\end{multline}
where we define 
\begin{equation}
	\Bar{T}_{iq,pa}(\tau_{34})=\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int \SO{a}(\bx_3) \SO{i}(\bx_4)\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34}) \psi_p(\bx_6)\psi^*_{q}(\bx_5)
\end{equation}
Using the definition $X_{qp,S}=\mel{N}{\hat{c}^\dagger_{q}\hat{c}_p}{N,S}$, we arrive at

\begin{equation}
	\qty(\e{a}{}-\e{i}{}-\omega_S) X_{ia,S}+\sum_{pq}X_{qp,S}\widetilde{T}_{iq,pa}(\omega_S)=0,
\end{equation}
where we define
\begin{equation}
	\widetilde{T}_{iq,pa}(\omega_S)
	= \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\Bar{T}_{iq,pa}(\omega)e^{-i\omega\eta} \left[\frac{1}{\omega_S-\omega+\e{q}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{p}{}-\e{a}{}+i\eta}\right] 
 \label{Eqn:H2preso_eigen_2_T}
\end{equation}

This is an eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as,
\begin{equation}
\label{eq:H2preso_eigen}
	H_{ia,qp}(\omega_S) X_{qp,S}=\omega_S X_{ia,S}
\end{equation}
with
\begin{equation}
	H_{ia,qp}(\omega_S)=(\e{a}{}-\e{i}{})\delta_{iq}\delta_{ap} + \mel{ip}{}{aq}+\widetilde{T}_{iq,pa}(\omega_S).
\end{equation}
Here $H_{ia,jb}(\omega_S)=A_{ia,jb}(\omega_S)$ and $H_{ia,bj}(\omega_S)=B_{ia,jb}(\omega_S)$.
If we drop the correlation part of $\Bar{T}$ [the first two terms on the right-hand side of Eq.~\eqref{eq:T_bar}] we obtain the TDHF equations. To calculate the correlation contribution we use \alert{Eq.~(9)} in Eq.~\eqref{Eqn:H2preso_eigen_2_T}, and after integration over the frequency we arrive at \alert{Eq.~(21).}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\twocolumngrid
\section*{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \hypertarget{SI}{data} that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.

%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{Tmatrix}
%%%%%%%%%%%%%%%%%%%%%%%%
